### 2019-01-18 supplement.Rmd
start of supplementay materials

---
title: 'Supplementary Materials'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  bookdown::word_document2:
    reference_docx: word-styles-reference-01.docx
bibliography: eyeIntegration_v1_app_manuscript.bib
csl: investigative-ophthalmology-and-visual-science.csl
---

```{r}
load('data/dimRed_plot_data.Rdata')
library(tidyverse)
library(ggrepel)

# eye PCA
eye_pca_plot <- ggplot(dimRed_plot_prep_eye_pca %>% 
         mutate(Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label)),
       aes(x=`Dimension 1`,y=`Dimension 2`)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  ggtitle(paste0('Eye Tissue PCA')) +
  geom_point(size=16, alpha=0.1, aes(colour=Tissue)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  geom_label_repel(aes(label=Label), alpha=0.8, size=4, box.padding = unit(0.3, "lines")) + theme_minimal() +
  xlab('Dimension 1') + ylab('Dimension 2')
# eye tSNE
eye_tsne_plot <- ggplot(dimRed_plot_prep_eye_tsne %>% 
                         mutate(Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label)),
                       aes(x=`Dimension 1`,y=`Dimension 2`)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  ggtitle(paste0('Eye Tissue t-SNE')) +
  geom_point(size=16, alpha=0.1, aes(colour=Tissue)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  geom_label_repel(aes(label=Label), alpha=0.8, size=4, box.padding = unit(0.3, "lines")) + theme_minimal()
plot_grid(eye_pca_plot, eye_tsne_plot)

# pca
pca_plot <- ggplot(dimRed_plot_prep_pca %>% 
                     mutate(Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label)),
                   aes(x=`Dimension 1`,y=`Dimension 2`)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  ggtitle(paste0('All Tissue PCA')) +
  geom_point(size=16, alpha=0.1, aes(colour=Tissue)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  geom_label_repel(aes(label=Label), alpha=0.8, size=4, box.padding = unit(0.3, "lines")) + theme_minimal() +
  xlab('Dimension 1') + ylab('Dimension 2')

# tSNE
tsne_plot <- ggplot(dimRed_plot_prep_tsne %>% 
                         mutate(Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label)),
                       aes(x=`Dimension 1`,y=`Dimension 2`)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  ggtitle(paste0('All Tissue PCA')) +
  geom_point(size=16, alpha=0.1, aes(colour=Tissue)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  geom_label_repel(aes(label=Label), alpha=0.8, size=4, box.padding = unit(0.3, "lines")) + theme_minimal() +
  xlab('Dimension 1') + ylab('Dimension 2')

```